---
title: "Miscellaneous Git Stuff"
author: "Jeffrey Arnold"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Miscellaneous code used to do stuff with the GitHub API for this course.

## Setup

```{r}
library("tidyverse")
library("glue")
library("gh")
```

GitHub organization of the course:
```{r}
ORG <- "UW-POLS501"
```
GitHub team for students in this quarter's course:
```{r}
STUDENT_TEAM <- "2018 students"

```
```{r}
INSTRUCTOR_TEAM <- "2018 instructors"
```
Main repository for the course:
```{r}
REPOS <- "2018"
```
Instructors usernames. This could also be queried with the API.
```{r}
INSTRUCTORS <- c("jrnold", "calvinhgarner") 
```


## Get Team

Get a team by name:
```{r}
get_team <- function(teamname) {
  gh("GET /orgs/:org/teams", org = ORG) %>%
    keep(~ .x$name == teamname) %>%
    pluck(1)
}
```

```{r}
instructor_team <- get_team(INSTRUCTOR_TEAM)
student_team <- get_team(STUDENT_TEAM)
```


## Issues

Create an issue for each assignment 1-9

```{r}
create_assignment_issue <- function(i) {
  gh("POST /repos/:owner/:repo/issues",
     owner = ORG, repo = REPOS,
     title = glue("Write assignment {i}"),
     body = glue("Write assignment {i}.",
                 "- [ ] Create submission template in repo ",
                 "- [ ] Write asssignment in content/assignments/assignment-{i}.md", 
                 .sep = "\n"))
}
walk(1:9, create_assignment_issue)
```

Create an issue for the solutions in assignments 1-9
```{r}
create_assignment_key_issue <- function(i) {
  gh("POST /repos/:owner/:repo/issues",
     owner = ORG, repo = REPOS,
     title = glue("Write solutions to assignment {i}"),
     body = glue("Solutions should go in the repository ",
                 "`UW-POLS501/assignment-{i}-solutions`.",
                 "Students should not have access to it until solutions are released."),
                 .sep = "\n")
}
walk(1:9, create_assignment_key_issue)
```

# Create Student Repositories and Teams

List of all students member info
```{r}
student_members <- gh("/teams/:id/members", id = student_team$id,
                           role = "member")
```

Extract their login names
```{r}
# get list of usernames
student_logins <- map_chr(student_members, "login")
```

Create a team for a student that includes only the student and instructors.
This team will have read/write access to their repo.
```{r}
create_student_team <- function(username) {
  team <- get_team(username)
  if (is.null(team)) {
    team <- 
      gh("POST /orgs/:org/teams", org = ORG, maintainers = INSTRUCTORS, 
         name = username, description = glue("Team for {username}'s project"),
         privacy = "closed")    
  } else {
    cat(glue("Team {username} already exists"))
  }
  gh("PUT /teams/:id/memberships/:username", id = team$id, 
      username = username, role = "member")

}

```
```{r}
map(student_logins, create_student_team)
```

For each student, create a repository :

- instructor's team has admin privileges
- student has push/pull privileges
- other students have pull privileges
- initialize repository it with a `README` and `.gitignore` file

```{r}
create_student_repo <- function(username) {
 repo <- gh("POST /orgs/:org/repos", 
            org = ORG,     
            name = username,
            desciption = glue("POLS/CSSS 501 (Winter 501) project of {username}"),
            private = TRUE,
            has_issues = TRUE,
            has_wiki = FALSE,
            auto_init = TRUE,
            gitignore_template = "R")
  gh("PUT /teams/:id/repos/:org/:repo", org = ORG,
     id = instructor_team$id, repo = username,
     permission = "admin")
  gh("PUT /teams/:id/repos/:org/:repo", org = ORG,
     id = get_team(username)$id,
     repo = username, permission = "push")
  gh("PUT /teams/:id/repos/:org/:repo", org = ORG,
     id = student_team$id, repo = username,
     permission = "pull")  
}

```

```{r}
# walk(student_logins, create_student_repo)
walk("nfullman", create_student_repo)
```



## Notes

Currently there is no nesting of the teams. In the future, it would make sense to nest
them by course.

## Scripting Assignment Submissions

It is probably better to have students manually do this, but it could be scripted.

```{r}
# Submit an assignment via the API. Creates a release and an issue
submit_assignment <- function(corrected = FALSE) {
  owner <- "UW-POLS501"
  repo <- "2018"
  assignees <- list("calvinhgarner", "jrnold")
  # Assumes master/head to get latest repo. Could also test by time.
  latest <- gh("/repos/:owner/:repo/commits",
               owner = owner, repo = repo)[[1]]

  # create a tag
  # this automatically creates a release
  tagname <- "submission"
  gh("POST /repos/:owner/:repo/git/refs",
            owner = owner, repo = repo,
            ref = glue("refs/tags/{tagname}"),
            sha = latest[["sha"]])

    # Create issue
   gh("POST /repos/:owner/:repo/issues",
      owner = owner, repo = repo,
      title = if (corrected) "Review Corrections" else "Review Submission",
      body = glue("Please review submission {latest$sha}."),
      assignees = assignees)
}

```

