---
title: "Untitled"
author: "Jeffrey Arnold"
date: "1/29/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loops in R

Read over the sections in *R for Data Science* on iteration and the 

It briefly mentions "\n" as newline character.


# General Conditional Statements in R

Recall that in general the `return` function is not needed in functions since a function implictly return the value of its last expression.
However, in some functions in which you use conditional statements you may need to return values *early*. 
The situation would look like this pseudo-code.
```{r}
foo <- function(x) {
  if (x == something) {
    # some calculations
    # return a value early
    return(...)
  }
  # other calculations
  ...
  # last value is returned 
}
```

# Poll Predictions

- How does prediction allow for black-boxing the method? 
- What is prediction error?
- What is unbiased?
- False (true) positive, false (true) negative

# Linear regression

- Correlation
- Least squares
- Residual or prediction error
- RMSE
- How are intercept and slope parameters calculated?
- What is the relationship between correlation and beta?
-


# Regression to the mean

What is the phenomena of regresssion to the mean? 

# Merging Data Sets in R

We've already covered that
- Review the output of `lm`
- What is the mean of errors? What is the median of errors?


