---
title: Supplementary Notes for *R for Data Science*
author: Jeffrey Arnold
editor_options: 
  chunk_output_type: console
---

Attached are

## Data Visualization

## Workflow: basics

## Data Transformation

## Exploratory Data Analysis

## Workflow: Projects

## Variable names in R

## Tibbles

- What can variables be named in R?
- What are non-syntactic variables?

## Data import

- What are csv files?
- How does a csv reader determine the types of variables?
- Read Stata and SPSS files

## Tidy data

Objectives

- Describe what "tidy data" frame is and why it is useful
- Describe when non-tidy data formats may be useful
- Convert data from long to wide formats using `spread()` and `gather()`
- Combine and split variables using `unite()` and `gather()`


This is an especially import chapter. Read it carefully.

### Common idioms 

There are a few common idioms that I use `spread()` and `gather()` for.

#### Summarizing multiple variables

Reshaping a data frame to a long format with columns for variables and values is a convenient format to calculate summary statistics for multiple variables.
```{r results='asis'}
library("tidyverse")
library("gapminder")

gapminder_summary <-
  gapminder %>%
  select(lifeExp, pop, gdpPercap) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(mean = mean(value), sd = sd(value))
gapminder_summary
```

If necessary, the data can be reshaped so each (variable, statistic) combination is a column:
```{r}
gapminder_summary %>%
  gather(stat, value, -variable) %>%
  # create column with columns variable, stat, value
  unite(variable_stat, variable, stat) %>%
  # combine variable and stats into a single column
  spread(variable_stat, value)
  # spread into a single column
```

An alternative method to summarize multiple variables is to use the `summarise_at` function:
```{r}
summarise_at(gapminder, vars(gdpPercap, lifeExp, pop), 
             funs(mean, sd))
```

#### Plotting multiple variable

If multiple variables need to be plotted, an alternative to creating multiple plots is to create a long data frame with a variable column and then facet by variable. 
For example, to plot the distribution of several variables in the same plot:
```{r}
library("tidyverse")
library("gapminder")

gapminder_long <-
  gapminder %>%
  select(lifeExp, pop, gdpPercap) %>%
  gather(variable, value)

ggplot(gapminder_long, aes(x = value)) +
  geom_density() +
  facet_wrap(~ variable, scales = "free", ncol = 1)
  scale_x_log10()
```


## Character vectors

- Understand what a character vector is and how to use functions in the string
- Understand what character encodings are and how to 
- Understand what escape sequences are and how to use them in R character sequences.
- Understand what regular expressions are and use them to

    - filter filenames returned by `list.files()`
    - match variable names in `select()` using the `match` function
    - filter by patterns in a character vector using `filter` and `str_detect`
    - replace patterns in strings using `str_replace`
    - extract information from a character vector using `str_match`
    
### Regular expressions

Match all filenames ending in `.csv` or `.tsv`
```{r}
filenames <- c("foo.csv", "foo.tsv", "foo.RData", "foo.dta", "tsv.doc",
               "csv_filenames.txt")
# to return a logical vector
str_detect(filenames, "\\.[ct]sv$")
# to return the matching elements 
str_subset(filenames, "\\.[ct]sv$")
```

Extract the name, party, and state from U.S. Congressperson names formatted according
to the AP Style:
```{r}
congress <- c("Ryan (R-WI)", "Pelosi (D-CA)", "McCarthy (R-CA)", 
              "Scalise (R-LA)", "Hoyer (D-MD)")
str_match(congress, "^(.*) \\(([A-Z])-([A-Z][A-Z])\\)")
```

Turn multiple whitespace values into only one space:
```{r}
str_replace_all(c("a a", "a  a", "a   \t    a    a a     "),
                "\\s+", " ")
```

Tokenize a text into "words":
```{r}
text <- readLines(system.file("extdata", "federalist", "fp01.txt", 
                              package = "qss")) %>%
  str_c(collapse = " ") %>%
  str_match_all("[A-Za-z]+") 

text <- tibble(text = text[[1]][, 1]) %>%
  mutate(text = str_to_lower(text)) %>%
  count(text) %>%
  arrange(desc(n))
```


### Encoding

Character encoding can occur when reading data from csv and text files.

Character encoding problems will occur at some point. It is becoming less of an issue due to a standardization to UTF-8, but it still occurs frequently.
The most likely encoding that you will encounter (if working with European texts) is Windows-1252 (CP-1252) or ISO-8859-1.

### Escape sequences

Certain characters preceeded by a backslash are treated specially. The most common are: `\n` (new-line), `\t` (tab).

Use a double-backslash `\\` for backslash in character vectors.

One common error is using a backslash in Windows file paths, which results in an error:
```{r error=TRUE}
"path\to\something"
```
Escaping the backslash will work, but won't work on MacOS or Linux.
It is better to either always use forward-slashes or `file.path` when writing paths in R.
```{r}
file.path("path", "so", "something")
file.path("path/to/something")
```


## Dates and Times

- Be able to convert vectors to and from strings
- Understand how dates are stored numerically. What is an epoch?
- What is UTC? 
- What is an impotant difference between using seconds, minutes, hours or days and
    using months and years as units of time? 

Date-time variables are almost always a pain to work with.
Common issues include the irregularity of months and time-zones.


## Factors

- Understand the difference between character vector and factor and when to use each

When to use factors?

    - You need to store metadata, e.g. levels
    - Control over the order of categories for plotting in ggplot
    - Models, e.g. `lm`

See the Stat 545 [discussion of factors](http://stat545.com/block029_factors.html).

## Lists and Vectors

- What are atomic vectors? What is the difference between lists and atomic vectors?
- What is `NULL` in R? How do you test for it?
- What is the difference between a matrix and a data frame? When should you use each?
- What are floating point numbers? What issues arise with them?
- Why are floating point numbers sometimes called "double" or use postfix `dbl` in tidyverse functions? 

When should you use lists? 
Lists can be used when you have several variables that you want to keep together, or will need to access programmatically in a loop or functionals.
Oftentimes I get the question, how can a programmatically name variables. 
This is possible, but almost always the answer is to create a list and save the values in different elements which are programmatically named.

