## Messy Data

- 
- Map COW datasets as relational datasets

Notable errors in Research

- https://bids.berkeley.edu/news/call-reproducibility-workflows
- http://www.peri.umass.edu/fileadmin/pdf/working_papers/working_papers_301-350/WP322.pdf
- https://genomebiology.biomedcentral.com/articles/10.1186/s13059-016-1044-7

Syllabus on reproducible rsearch: https://github.com/psu-psychology/psy-511-reproducible-research-spring-2017/blob/master/lecture-notes/2017-04-10-whole-elephant.md

http://ropensci.github.io/reproducibility-guide/sections/references/

REcent

- "Evaluating replicability of laboratory experiments in economics" 10.1126/science.aaf0918
- "Is Economics Research Replicable? Sixty Published Papers from Thirteen Journals Say ”Usually Not”" https://www.federalreserve.gov/econresdata/feds/2015/files/2015083pap.pdf
- https://thepoliticalmethodologist.com/2014/12/09/a-decade-of-replications-lessons-from-the-quarterly-journal-of-political-science/


## Types of Questions

- Description
- Prediction
- Causation


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Resources

Simpson's Paradox

```{r}

```

- http://homepage.stat.uiowa.edu/~mbognar/1030/Bickel-Berkeley.pdf
- World Income Inequality. http://www.nber.org/papers/w8904
- Income and Partisanship (traditionally) 
- [US Median Wage Decline](http://blog.revolutionanalytics.com/2013/07/a-great-example-of-simpsons-paradox.html)
- Norris, Floyd. "[Can Every Group Be Worse Than Average? Yes](https://economix.blogs.nytimes.com/2013/05/01/can-every-group-be-worse-than-average-yes/)"
- [Beware the Lurking Variable](http://www.statlit.org/pdf/2006SchieldSTATS.pdf) has some clear examples.
- https://www.ft.com/content/94e3acec-a767-11e7-ab55-27219df83c97
- https://ww2.amstat.org/publications/jse/secure/v7n3/datasets.morrell.cfm

It is not necessarily causal. 

Hoynes, Hilary W., et al. “Poverty in America: Trends and Explanations.” The Journal of Economic Perspectives, vol. 20, no. 1, 2006, pp. 47–68. JSTOR, JSTOR, www.jstor.org/stable/30033633.
