---
title: "Installing R"
---

## Installing or Updating R

### Install R

The R website is ... special. Since you probably don't know or care what a mirror is, here are links to the

- Go to the [R Project](https://cran.r-project.org/) website
- Choose the `Download and install R` your operation system:
    - Mac: Download `R-3.4.3.pkg`
    - Windows: First click on `base` and then on `Download R 3.4.3 for Windows`

### Upgrade R

If you already have R installed, to check that you have the newest version of R, run
```{r eval=FALSE}
sessionInfo()
```
which will produce output like,
```
R version 3.4.3 (2016-03-10)
Platform: x86_64-apple-darwin13.4.0 (64-bit)
Running under: OS X 10.11.3 (El Capitan)

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] R6_2.1.2

loaded via a namespace (and not attached):
[1] rsconnect_0.4.2 tools_3.4.3     knitr_1.12.3
```
At the start of this course, January 5, 2018, the newest version of R is 3.4.3.
Check that your `R version` is that one.

### Update packages

Now, update all your packages
```{r eval=FALSE}
update.packages(ask = FALSE)
```
The option `ask = FALSE` answers "yes to all the questions" as to whether to answer


## Installing or Upgrading R RStudio

If you don't have RStudio installed, install it from [rstudio.com](https://www.rstudio.com/products/RStudio/#Desktop). You want to download the RStudio **Desktop Open Source Edition** (which is FREE).

If you already have RStudio installed, go to `Help > Check for Updates` in the menu and install a new version if the pop-up window indicates that there is a new version available.

## Installing Tools to Build R Packages from Scratch

Mature R packages are distributed over CRAN and installable in R using `install.packages` without any other assistance.
However, we may use some bleeding-edge R packages that require the packages to be built from source and installed using the [devtools](http://www.rstudio.com/products/rpackages/devtools/) package.
```{r eval=FALSE}
install.packages("devtools")
```

This requires some extra [programs](http://www.rstudio.com/products/rpackages/devtools/).

- Windows: Install the newest version of [Rtools](http://cran.r-project.org/bin/windows/Rtools/). Use default settings.
- Mac: Install Xcode from the Mac App Store.

To check that this is setup correctly, run the following
```{r eval=FALSE}
library("devtools")
has_devel()
```
If everything is correctly set up, it will return `TRUE`.
If you do not have Rtools or Xcode installed you will get an error message similar to this:
```
> has_devel()
"C:/PROGRA~1/R/R-32~1.4RE/bin/x64/R"  \
  --no-site-file --no-environ --no-save  \
  --no-restore CMD SHLIB foo.c

Warning message:
running command 'make -f "C:/PROGRA~1/R/R-32~1.4RE/etc/x64/Makeconf" -f "C:/PROGRA~1/R/R-32~1.4RE/share/make/winshlib.mk" SHLIB="foo.dll" WIN=64 TCLBIN=64 OBJECTS="foo.o"' had status 127
Error: Command failed (1)
```


## Installing LaTeX

LaTeX is needed to create PDFs from R Markdown files.
Install the following, depending on your operating system:

- Windows: [MikTex](http://miktex.org/download)

    - When asked choose default paper size of "letter" instead of "A4"

- MacOS: [MacTex](https://tug.org/mactex/)

Note, these are very large files and can take a while to download.
